The effects of varying depth in cosmic shear surveys

### 2019-10-01 paperdraft.tex
added marikas changes

%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
%                              according to the author-year natbib style
%
\documentclass{aa}  
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\ben}{\begin{enumerate}}
\newcommand{\een}{\end{enumerate}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\bmi}[1]{\begin{minipage}{#1 cm}}
\newcommand{\emi}{\end{minipage}}
\renewcommand{\[}{\begin{equation}}
\renewcommand{\]}{\end{equation}}

@@

   \author{Sven Heydenreich\inst{1}%\fnmsep\thanks{sven@astro.uni-bonn.de}
          \and Peter Schneider\inst{1} 
          \and Hendrik Hildebrandt\inst{2,1}
          \and Catherine Heymans\inst{3,2} 
          \and Marika Asgari\inst{3}
          \and Jan Luca van den Busch\inst{2,1}
          \and Chieh-An Lin\inst{3,5,6}
          \and Benjamin Joachimi\inst{4}
          \and Tilman Tr\"{o}ster\inst{3}
          }
   \institute{Argelander-Institut f\"ur Astronomie, Auf dem H\"ugel 71, 53121 Bonn, Germany 
\and 
Astronomisches Institut, Ruhr-Universit\"at Bochum, Universit\"atsstr. 150, 44801 Bochum, Germany 
\and
Institute for Astronomy, University of Edinburgh, Royal Observatory, Blackford Hill, Edinburgh EH9 3HJ, UK
\and
Department of Physics and Astronomy, University College London, Gower Street, London WC1E 6BT, UK
\and
Service d'Astrophysique, CEA Saclay, Orme des Merisiers, B\^at 709, 91191 Gif-sur-Yvette, France
\and 
Fenglin Veteran Hospital, 2 Zhongzheng Rd. Sec. 1, Fenglin Township, Hualien 97544, Taiwan
\\ \email{sven@astro.uni-bonn.de, peter@astro.uni-bonn.de}
             }

@@

  \abstract
% context (optional)
   {Cosmic shear proves to be a powerful tool to study the properties of the Universe. The discrepancy in the parameter $S_8 = \sigma_8\sqrt{\Omega_{\rm{m}}/0.3}$ between measurements in the local Universe and the cosmic microwave background motivates further investigation of yet unaccounted systematic biases. 
% aims
   Ground-based surveys are subject to a variation in depth caused by varying atmospherical conditions; we investigate and quantify the resulting effects. In particular, we check whether they introduce a bias to the resulting cosmological parameters and if they can be responsible for the occurrence of B-modes.
% methods
   We construct a semi-analytic model to estimate the impact on the shear correlation functions and analyse the implications for cosmological parameters. Furthermore, we construct COSEBIs of the correlation functions to quantify the occurring B-modes.
% results
   For the Kilo-Degree Survey this effect introduces an error in $\xi_\pm$ of the order of a few percent on small scales. We find that the resulting bias in $\Omega_{\rm m}$ and $\sigma_8$ will be significant in next-generation cosmic shear surveys, but that the parameter $S_8$ is surprisingly robust against this modification. We find that the dependence of this effect on the underlying cosmology is not negligible. We also report the occurrence of B-modes, although at a not yet significant level.
% conclusions (optional)
   We conclude that the effects of varying depth for current cosmic shear surveys on cosmological parameters are not yet significant, but should be accounted for in next-generation experiments.}

@@

%
%-------------------------------------------------------------------
\section{Introduction}
The discovery of cosmic shear has provided us with a new and powerful cosmological tool to empirically test the standard model of cosmology and to determine its parameters. Contrary to the analysis of the cosmic microwave background \citet[CMB, e.g.][]{2018arXiv180706209P}, cosmic shear is more sensitive to the properties of the local Universe and thus provides an excellent consistency check for the standard model. Current cosmic shear surveys are particularly sensitive to the parameter $S_8=\sigma_8 \sqrt{\Omega_{\rm m}/0.3}$, where $\sigma_8$ characterizes the normalization of the matter power spectrum and $\Omega_{\rm m}$ is the matter density parameter.
It is interesting to note that all three current major cosmic shear results report a lower $S_8$ than inferred from the CMB analysis by \citet{2018arXiv180706209P}. While their analysis determined a value of $S_8 = 0.830 \pm 0.013$, \citet{2018arXiv180909148H} report $S_8 = 0.800^{+0.029}_{-0.028}$ from an analysis of the Subaru Hyper Suprime-Cam survey, \citet[][hereafter H18]{2018arXiv181206076H} obtained $0.737_{-0.036}^{+0.040}$ from KiDS+VIKING data, and the Dark Energy Survey \citep[DES,][]{2018PhRvD..98d3528T} arrived at $S_8=0.782\pm 0.027$. Also, \citet{2013MNRAS.432.2433H} report $S_8 = 0.759 \pm 0.020$ from their analysis of CFHTLens data. After reweighting the obtained redshift of data from DES and combining this with KiDS data \citep{Joudaki:2019} yield even higher tension with the value determined for $S_8$ by the Planck Collaboration. This discrepancy has received a lot of attention \citep{2017MNRAS.471.1259J,2016MNRAS.459..971K}. It could be interpreted as a statistical coincidence, a sign of new physics like massive neutrinos \citep{2014PhRvL.112e1303B}, time-varying dark energy or modified gravity \citep{2016A&A...594A..14P}; or as the manifestation of a systematic effect, either in the cosmic shear surveys or in the Planck mission \citep{2016ApJ...818..132A}, that is not yet accounted for.

@@

The depth of a pointing primarily influences the number density as well as the average redshift of galaxies, meaning that a variation in depth is responsible for a correlation between those two quantities (see Fig. \ref{fig:nz_of_meanz}).
%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{images/cov_nz_meanz.pdf}
\caption{Weighted number of galaxies $N$ and average redshift $\la z\ra$ in the KiDS+VIKING-450 survey \citep[KV450,][]{Wright:2018} in pointings of different depth for each of the five tomographic bins used in H18. Each colour corresponds to one redshift bin of H18, the fainter points denote pointings of shallower depth. Details on how the data were obtained are described in Sect. \ref{sec:xipm_semianalytic}.}
\label{fig:nz_of_meanz}
\end{figure}
%
In Sect. \ref{sec:analytic_models} we will introduce two simple toy models to understand this effect and analyze the impact on the cosmic shear power spectrum. In Sect. \ref{sec:xipm_semianalytic} we will estimate the effect on the shear correlation functions $\xi_\pm$ using a semi-analytic model. We will present our results in Sect. \ref{sec:results}. In Sect. \ref{sec:discussion} we will discuss our results and comment on the impact of our used simplifications. We will assume the standard weak gravitational lensing formalism, a summary of which can be found in \citet{2001PhR...340..291B}.

%-------------------------------------------------------------------
\section{Simple, analytic toy models}
\label{sec:analytic_models}
For our first analysis we assume that all the matter between sources and observer is concentrated in a single lens plane of distance $D_{\rm d}$ from the observer. If we now distribute sources at varying distances $D_{\rm s}$, then the convergence $\kappa$ varies according to $\kappa \propto D_{\rm{ds}}/D_{\rm s}$, where $D_{\rm{ds}}$ is the distance between the lens plane and the respective source.

Assuming that the depth, and thus the source redshift population, varies between pointings of the camera, an observer will measure a shear signal that is modified by a step-like depth-function, $\gammao(\b\theta)=W(\b \theta)\gamma(\b\theta)$, where $W$ is proportional to the mean of the lensing efficiency $\Dds/\Ds$ of one pointing and $\gamma$ denotes the shear that this pointing would read if it were of the average depth.
We can parametrize $W$ as $W(\bm{\theta}) = 1+w(\bm{\theta})$, where $\langle w(\bm{\theta})\rangle=0$ holds. 
\subsection{Modelling the power spectrum}
\label{sec:modelling_power_spectrum}
In our first model we describe the impact of varying depth on the power spectrum, following the simplifications described above.
% Following \citet{1993ApJ...404..441K}, we can simulate the shear $\gamma$ from the convergence by convolution with the kernel \[
%\mathcal{D}(\bm{\theta}) = \frac{\theta_2^2-\theta_1^2-2i\theta_1\theta_2}{\pi |\bm{\theta}|^4}\, .
%\]

%Similarly, an observer that measures the shear $\gamma$ can reconstruct the convergence $\kappa$ by convolution with the kernel $\mathcal{D}^*$, where ${}^*$ denotes the complex conjugate. Assuming that the optical depth, and thus the source redshift populations, vary between pointing, an observer will measure a shear-signal that is modified by a depth-function $W(\bm{\theta}) = 1-w(\bm{\theta})$, where $\langle w(\bm{\theta})\rangle=0$ holds. The reconstruction of the observed convergence $\kappa^{\rm obs}$ will thus not yield the original convergence. In particular there is no reason why the convolution with a complex kernel should have a nonzero imaginary part. This allows us to separate our received signal into E- and B-modes following \[
% \kappa = \kappa^E+i\kappa^B\, ,
%\]
%where $B$-Modes can not be generated by a lensing signal and are thus a tracer of remaining systematics.
%2002A&A...389..729S SRC

In accordance to the definition of the shear power spectrum \begin{equation}
\la \hat{\gamma}(\b\ell)\,\hat{\gamma}^*(\b\ell')\ra = (2\pi)^2\delta(\b\ell-\b\ell')P(|\b\ell|) \, ,
\label{eq:original_power_spectrum}
\end{equation}
where $\hat{\gamma}$ denotes the Fourier transform of $\gamma$, we define the observed power spectrum via \begin{equation}
P^{\text{obs}}(\b\ell) \equiv \frac{1}{(2\pi)^2}\int \d^2 \ell'\,  \la \gammaoh(\b \ell)\, \gammaoh {}^*(\b \ell')\ra \, .
\end{equation}
Note that due to the depth-function, both the assumptions of homogeneity and isotropy break down, which means that we can neither assume isotropy in the power spectrum, nor can we assume that $\la \gammaoh(\b \ell) \gammaoh {}^*(\b \ell')\ra$ vanishes for $\b\ell\neq\b\ell'$.
To model a constant depth on each individual pointing $\b \alpha$, we can choose random variables $w_{\b \alpha}$, that only need to satisfy $\langle w_{\b \alpha}\rangle=0$, and parametrize $w(\b\theta)$ as 
\begin{align}
w(\b \theta) = & \sum_{\b \alpha \in \mathbb{Z}^2} w_{\b \alpha} \Xi(\b \theta-L\b \alpha)\text{ , with the box-function } \\
\Xi(\b \theta) = & \begin{cases}
1 \qquad \b \theta\in \left[-\frac{L}{2},\frac{L}{2}\right]^2 \\
0 \qquad \text{else}
\end{cases},
\label{eq:defweightf}
\end{align}
where $L$ is the sidelength of one pointing.
Following the calculations in Appendix \ref{sec:calc of PS}, we derive 
\begin{equation}
P^{\text{obs}}(\b \ell)  =  P(\b \ell) + \la w^2\ra \int\frac{\text{d}^2\b k}{(2\pi)^2}\,\hat{\Xi}(\b \ell-\b k)\, P(\b k)\, ,
\end{equation}
where $\la w^2\ra \equiv \la w_{\b \alpha}^2\ra$ is the dispersion of the depth-function and $\hat{\Xi}$ is the Fourier transform of a box-function, which is a 2-dimensional sinc-function (compare \ref{sec:calc of PS}).
The observed power spectrum $P^{\text{obs}}$ is thus composed of the original power spectrum $P(\ell)$ from Eq. \eqref{eq:original_power_spectrum}, plus a convolution of the power spectrum with a sinc-function, scaling with the variance of the function $w(\b\theta)$. In particular, the power spectrum is not isotropic anymore.

@@

\subsection{Modelling the shear correlation functions}
\label{sec:xipm_analytic}
More convenient measures to infer cosmological information from observational data are the shear correlation functions $\xi_\pm$, which are defined as \[
\xi_\pm(\theta) = \la \gamma_{\rm t}\gamma_{\rm t}\ra(\theta) \pm \la \gamma_\times\gamma_\times\ra(\theta) \, .
\]
Here, $\gamma_{\rm{t}}$ and $\gamma_\times$ denote the tangential- and cross-component of the shear for a galaxy pair with respect to their relative orientation \citep[see][]{2002A&A...396....1S}.
The shear correlation functions are the prime estimators to quantify a cosmic-shear signal, since it is simple to include a weighting of the shear measurements into the correlation functions and, contrary to the power spectrum, one does not have to worry about the shape of the survey footprint or masked regions, or model the noise contribution. 
For this analysis we will follow the assumption that a deeper pointing shows a stronger shear signal $\gamma^{\rm{obs}}(\b\theta) = W(\b\theta)\gamma(\b\theta)$ as described above. While this is not true for a 3-dimensional lensing matter distribution, it should be a reasonable first approximation for small variations in mean source redshift. Additionally, we assume that a higher depth does not only lead to a stronger average shear, but also to a higher galaxy number density, implying a correlation between those two quantities.

Let $N^i(\b \theta),N^j(\b\theta)$ be the average weighted number of galaxies per pointing in redshift bins $i$ and $j$ and let $W^i(\b \theta),W^j(\b\theta)$ be the weighting of average shear. The observed correlation functions $\xi^{ij,\text{obs}}_\pm(\theta)$ now change from one of uniform depth, $\xi_\pm^{ij,\rm{uni}}(\theta)$, via \citep[see][]{2002A&A...396....1S}
\begin{align}
\xi^{ij,\text{obs}}_\pm(\theta) = & \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_{\rm t}(\bth)\gamma^{j,\rm{obs}}_{\rm t}(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \\
 & \pm \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_\times(\bth)\gamma^{j,\rm{obs}}_\times(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \nonumber\\
 = & \frac{\la N^i(\bth)N^j(\bthp)W^i(\bth)W^j(\bthp)\ra}{\la N^i(\bth)N^j(\bthp)\ra} \xi_{\pm}^{ij,\rm{uni}}(\theta) \, ,
 \label{eq:xipmblub1}
 \end{align}
 where the average $\la\ldots\ra$ represents both an ensemble average as well as an average over the position $\bth$.
 Assuming that the depth of different pointings is uncorrelated implies that the same holds for the functions $W$ and $N$. This means that the only important property of a galaxy pair is whether or not they lie in the same pointing. We want to denote the probability that a random galaxy pair of separation $\theta$ lies in the same pointing by $E(\theta)$. This function is depicted in Fig. \ref{fig:eoftheta_lin}, and an analytical expression is derived in App. \ref{sec:model_e}.

@@

\section{A semi-analytic model}
\label{sec:xipm_semianalytic}
The previously derived analytic model describes how varying depth between pointings modifies the correlation function due to the correlation between number density and average redshift of source galaxies. While this model serves as an intuitive first approximation, it completely ignores any effects from the large scale structure (LSS) between the closest and the most distant galaxy. Therefore, we do not expect this model to yield accurate, quantitative results for cosmic shear surveys.

@@

An inspection of KiDS-data showed that the redshift distribution of sources is highly correlated with the depth in the $r$-band (compare Fig. \ref{fig:nz_of_meanz}). We thus chose to separate the survey into 10 percentiles, sorted by $r$-band depth, i.e. if a pointing had a shallower depth than 90\% of the other pointings, it would belong to the first percentile, and so on. For each percentile $m$ and each tomographic redshift bin $i$ we can extract a weighted number of galaxies $N^i_m$ and a source redshift distribution $p^i_m(z)$ following the DIR method of H18\footnote{Instead of using the actual number of galaxies, we take the sum of their lensing weights, following H17. Due to this we account for different weighting of galaxies in the shear correlation functions as well as in the average redshift distribution.}. The weighted numbers of Galaxies can be seen in Fig. \ref{fig:nz_of_meanz}, whereas the source redshift distributions are depicted in Fig. \ref{fig:nofz_percentiles}.
%
\begin{figure}
\centering
\includegraphics[trim = {1cm 0 1.2cm 0.1cm}, clip, width=\linewidth]{images/dist_z_all.pdf}
\caption{The source redshift distributions $p^i_m(z)$ for the best, average, and worst percentile in each tomographic redshift bin.}
\label{fig:nofz_percentiles}
\end{figure}
% Asserting an arbitrary cosmological model, we can transform the source redshift distributions $p^i_m(z)$ into comoving distance probability distributions $\mathcal{L}^i_m(\chi)$.

>>>>

\documentclass{aa}

@@

   \author{Sven Heydenreich\inst{1}%\fnmsep\thanks{sven@astro.uni-bonn.de}
          \and Peter Schneider\inst{1} 
          \and Hendrik Hildebrandt\inst{2,1}
          \and Catherine Heymans\inst{3,2} 
          \and Marika Asgari\inst{3}
          \and Jan Luca van den Busch\inst{2,1}
          \and Chieh-An Lin\inst{3}
          \and Benjamin Joachimi\inst{4}
          \and Tilman Tr\"{o}ster\inst{3}
          }
   \institute{Argelander-Institut f\"ur Astronomie, Auf dem H\"ugel 71, 53121 Bonn, Germany 
\and 
Astronomisches Institut, Ruhr-Universit\"at Bochum, Universit\"atsstr. 150, 44801 Bochum, Germany 
\and
Institute for Astronomy, University of Edinburgh, Royal Observatory, Blackford Hill, Edinburgh EH9 3HJ, UK
\and
Department of Physics and Astronomy, University College London, Gower Street, London WC1E 6BT, UK
% \and
% Service d'Astrophysique, CEA Saclay, Orme des Merisiers, B\^at 709, 91191 Gif-sur-Yvette, France
% \and 
% Fenglin Veteran Hospital, 2 Zhongzheng Rd. Sec. 1, Fenglin Township, Hualien 97544, Taiwan
\\ \email{sven@astro.uni-bonn.de, peter@astro.uni-bonn.de}
             }

@@

  \abstract
% context (optional)
   {Cosmic shear proves to be a powerful tool to study the properties of the Universe. The discrepancy in the parameter $S_8 = \sigma_8\sqrt{\Omega_{\rm{m}}/0.3}$ between measurements in the local Universe and the cosmic microwave background motivates further investigation of yet unaccounted systematic biases. 
% aims
   Ground-based surveys are subject to a variation in depth caused by varying atmospheric conditions; we investigate and quantify the resulting effects. In particular, we check whether they introduce a bias to the resulting cosmological parameters and if they can be responsible for the occurrence of B-modes.
% methods
   We construct a semi-analytic model to estimate the impact on the shear correlation functions and analyse the implications for cosmological parameters. Furthermore, we construct COSEBIs from the correlation functions to quantify the occurring B-modes.
% results
   For the Kilo-Degree Survey this effect introduces an error in $\xi_\pm$ of the order of a few percent on small scales. We find that the resulting bias in $\Omega_{\rm m}$ and $\sigma_8$ will be significant in next-generation cosmic shear surveys, but that the parameter $S_8$ is surprisingly robust against this modification. We find that the dependence of this effect on the underlying cosmology is not negligible. We also report the occurrence of B-modes, although at a not yet significant level.
% conclusions (optional)
   We conclude that the effects of varying depth for current cosmic shear surveys on cosmological parameters are not yet significant, but should be accounted for in next-generation experiments.}

@@

%
%-------------------------------------------------------------------
\section{Introduction}
The discovery of cosmic shear has provided us with a new and powerful cosmological tool to empirically test the standard model of cosmology and to determine its parameters. Contrary to the analysis of the cosmic microwave background \citep[CMB, e.g.~by][]{2018arXiv180706209P}, cosmic shear is more sensitive to the properties of the local Universe and thus provides an excellent consistency check for the standard model. Current cosmic shear surveys are particularly sensitive to the parameter $S_8=\sigma_8 (\Omega_{\rm m}/0.3)^\alpha$, where $\sigma_8$ characterizes the normalization of the matter power spectrum and $\Omega_{\rm m}$ is the matter density parameter. The parameter $\alpha\approx 0.5$ depends on the exact survey properties; in this paper we will fix it to be $\alpha = 0.5$.
It is interesting to note that all three current major cosmic shear results report a lower $S_8$ than inferred from the CMB analysis by \citet{2018arXiv180706209P}. While their analysis determined a value of $S_8 = 0.830 \pm 0.013$, \citet{2018arXiv180909148H} report $S_8 = 0.800^{+0.029}_{-0.028}$ from an analysis of the Subaru Hyper Suprime-Cam survey, \citet[][hereafter H18]{2018arXiv181206076H} obtained $0.737_{-0.036}^{+0.040}$ from KiDS+VIKING data, and the Dark Energy Survey \citep[DES,][]{2018PhRvD..98d3528T} arrived at $S_8=0.782\pm 0.027$. Also, \citet{2013MNRAS.432.2433H} report $S_8 = 0.759 \pm 0.020$ from their analysis of CFHTLens data. After reweighting the obtained redshift of data from DES and combining this with KiDS data \citep[][Asgari et al.~in prep.]{Joudaki:2019} yield even higher tension with the value determined for $S_8$ by the Planck Collaboration. This discrepancy has received a lot of attention \citep{2017MNRAS.471.1259J,2019arXiv190103686S,2019arXiv190911006T}. It could be interpreted as a statistical coincidence, a sign of new physics like massive neutrinos \citep{2014PhRvL.112e1303B}, time-varying dark energy or modified gravity \citep{2016A&A...594A..14P}; or as the manifestation of a systematic effect, either in the cosmic shear surveys or in the Planck mission \citep{2016ApJ...818..132A}, that is not yet accounted for.

@@

%The depth of a pointing primarily influences the number density as well as the average redshift of galaxies, meaning that a variation in depth is responsible for a correlation between those two quantities (see Fig. \ref{fig:nz_of_meanz}).
%

In Sect. \ref{sec:analytic_models} we will introduce two simple toy models to understand this effect and analyze the impact on the cosmic shear power spectrum. In Sect. \ref{sec:xipm_semianalytic} we will estimate the effect on the shear correlation functions $\xi_\pm$ using a semi-analytic model. We will present our results in Sect. \ref{sec:results}. In Sect. \ref{sec:discussion} we will discuss our results and comment on the impact of our used simplifications. We will assume the standard weak gravitational lensing formalism, a summary of which can be found in \citet{2001PhR...340..291B}.

%-------------------------------------------------------------------
\section{Simple, analytic toy models}
\label{sec:analytic_models}
For our first analysis we assume that all the matter between sources and observer is concentrated in a single lens plane of distance $D_{\rm d}$ from the observer. If we now distribute sources at varying distances $D_{\rm s}$, two effects become apparent: Firstly, the lensing efficiency $D_{\rm{ds}}/D_{\rm s}$ varies, where $D_{\rm{ds}}$ is the distance between the lens plane and the respective source. Secondly, and more importantly, for a more distant source \emph{more} matter is concentrated between the source and the observer, leading to a stronger shear signal.

Assuming that the depth, and thus the source redshift population, only varies between pointings of the camera, an observer will measure a shear signal that is modified by a step-like depth-function, $\gammao(\b\theta)=W(\b \theta)\gamma(\b\theta)$, where $W$ is proportional to the mean of the lensing efficiency $\Dds/\Ds$ of one pointing and $\gamma$ denotes the shear that this pointing would read if it were of the average depth.
We can parametrize $W$ as $W(\bm{\theta}) = 1+w(\bm{\theta})$. This implies that $\langle w(\bm{\theta})\rangle=0$ holds, where $\langle \cdot\rangle$ denotes the average over all pointings.
\subsection{Modelling the power spectrum}
\label{sec:modelling_power_spectrum}
In our first model we describe the impact of varying depth on the power spectrum, following the simplifications described above.
% Following \citet{1993ApJ...404..441K}, we can simulate the shear $\gamma$ from the convergence by convolution with the kernel \begin{equation}
%\mathcal{D}(\bm{\theta}) = \frac{\theta_2^2-\theta_1^2-2i\theta_1\theta_2}{\pi |\bm{\theta}|^4}\, .
%\end{equation}

%Similarly, an observer that measures the shear $\gamma$ can reconstruct the convergence $\kappa$ by convolution with the kernel $\mathcal{D}^*$, where ${}^*$ denotes the complex conjugate. Assuming that the optical depth, and thus the source redshift populations, vary between pointing, an observer will measure a shear-signal that is modified by a depth-function $W(\bm{\theta}) = 1-w(\bm{\theta})$, where $\langle w(\bm{\theta})\rangle=0$ holds. The reconstruction of the observed convergence $\kappa^{\rm obs}$ will thus not yield the original convergence. In particular there is no reason why the convolution with a complex kernel should have a nonzero imaginary part. This allows us to separate our received signal into E- and B-modes following \begin{equation}
% \kappa = \kappa^E+i\kappa^B\, ,
%\end{equation}
%where $B$-Modes can not be generated by a lensing signal and are thus a tracer of remaining systematics.
%2002A&A...389..729S SRC

In accordance to the definition of the shear power spectrum \begin{equation}
\la \hat{\gamma}(\b\ell)\,\hat{\gamma}^*(\b\ell')\ra = (2\pi)^2\delta(\b\ell-\b\ell')P(|\b\ell|) \, ,
\label{eq:original_power_spectrum}
\end{equation}
where $\hat{\gamma}$ denotes the Fourier transform of $\gamma$, we define the observed power spectrum via \begin{equation}
P^{\text{obs}}(\b\ell) \equiv \frac{1}{(2\pi)^2}\int \d^2 \ell'\,  \la \gammaoh(\b \ell)\, \gammaoh {}^*(\b \ell')\ra \, .
\end{equation}
Note that due to the depth-function, both the assumptions of homogeneity and isotropy break down, which means that we can neither assume isotropy in the power spectrum, nor can we assume that $\la \gammaoh(\b \ell) \gammaoh {}^*(\b \ell')\ra$ vanishes for $\b\ell\neq\b\ell'$.
To model a constant depth on each individual pointing, $\b \alpha$, we can choose random variables, $w_{\b \alpha}$, that only need to satisfy $\langle w_{\b \alpha}\rangle=0$. As we assume an infinite number of pointings, $\b\alpha$ can assume any two-dimensional integer value $\mathbb{Z}^2$ and we can parametrize $w(\b\theta)$ as 
\begin{equation}
w(\b \theta) = \sum_{\b \alpha \in \mathbb{Z}^2} w_{\b \alpha} \Xi(\b \theta-L\b \alpha)\, ,
\end{equation}
with the box-function
\begin{equation}
\Xi(\b \theta) = \begin{cases}
1 \qquad \b \theta\in \left[-\frac{L}{2},\frac{L}{2}\right]^2 \\
0 \qquad \text{else}
\end{cases},
\label{eq:defweightf}
\end{equation}
where $L$ is the sidelength of one pointing.
Following the calculations in Appendix \ref{sec:calc of PS}, we derive 
\begin{equation}
P^{\text{obs}}(\b \ell)  =  P(\b \ell) + \la w^2\ra \int\frac{\text{d}^2\b \ell'}{(2\pi)^2}\,\hat{\Xi}(\b \ell-\b \ell')\, P(\b \ell')\, .
\end{equation}
Here we have denoted $\la w^2\ra \equiv \la w_{\b \alpha}^2\ra$ as the dispersion of the depth-function, since the statistical properties of this function do not depend on the pointing $\b\alpha$. The Fourier transform of the box function, $\hat{\Xi}$, is a 2-dimensional sinc-function (see \ref{sec:calc of PS}).
The observed power spectrum, $P^{\text{obs}}$, is thus composed of the original power spectrum $P(\ell)$ from Eq. \eqref{eq:original_power_spectrum}, plus a convolution of the power spectrum with a sinc-function, scaling with the variance of the function $w(\b\theta)$.

@@

\subsection{Modelling the shear correlation functions}
\label{sec:xipm_analytic}
More convenient measures to infer cosmological information from observational data are the shear correlation functions $\xi_\pm$, which are defined as \begin{equation}
\xi_\pm(\theta) = \la \gamma_{\rm t}\gamma_{\rm t}\ra(\theta) \pm \la \gamma_\times\gamma_\times\ra(\theta) \, .
\end{equation}
Here, $\gamma_{\rm{t}}$ and $\gamma_\times$ denote the tangential- and cross-component of the shear for a galaxy pair with respect to their relative orientation \citep[see][]{2002A&A...396....1S}.
The shear correlation functions are the prime estimators to quantify a cosmic-shear signal, since it is simple to include a weighting of the shear measurements into the correlation functions and, contrary to the power spectrum, one does not have to worry about the shape of the survey footprint or masked regions, or model the noise contribution. 
For this analysis we will follow the assumption that a deeper pointing shows a stronger shear signal $\gamma^{\rm{obs}}(\b\theta) = W(\b\theta)\gamma(\b\theta)$ as described above. While this is not true for a 3-dimensional lensing matter distribution, it should be a reasonable first approximation for small variations in mean source redshift. Additionally, we assume that a higher depth does not only lead to a stronger average shear, but also to a higher galaxy number density, implying a correlation between those two quantities.

Let $N^i(\b \theta),N^j(\b\theta)$ be the average weighted number of galaxies per pointing in redshift bins $i$ and $j$ and let $W^i(\b \theta),W^j(\b\theta)$ be the weighting of average shear. The observed correlation functions $\xi^{ij,\text{obs}}_\pm(\theta)$ now change from uniform depth, $\xi_\pm^{ij,\rm{uni}}(\theta)$, via %\citep[see][]{2002A&A...396....1S}
\begin{align}
\xi^{ij,\text{obs}}_\pm(\theta) = & \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_{\rm t}(\bth)\gamma^{j,\rm{obs}}_{\rm t}(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \\
 & \pm \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_\times(\bth)\gamma^{j,\rm{obs}}_\times(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \nonumber\\
 = & \frac{\la N^i(\bth)N^j(\bthp)W^i(\bth)W^j(\bthp)\ra}{\la N^i(\bth)N^j(\bthp)\ra} \xi_{\pm}^{ij,\rm{uni}}(\theta) \, ,
 \label{eq:xipmblub1}
 \end{align}
 where the average $\la\ldots\ra$ represents both an ensemble average as well as an average over the position $\bth$.
 Assuming that the depth of different pointings is uncorrelated, the only important property of a galaxy pair is whether or not they lie in the same pointing. We want to denote the probability that a random galaxy pair of separation $\theta$ lies in the same pointing by $E(\theta)$. This function is depicted in Fig. \ref{fig:eoftheta_lin}, and an analytical expression is derived in App. \ref{sec:model_e}.

@@

\section{A semi-analytic model}
\label{sec:xipm_semianalytic}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{images/cov_nz_meanz.pdf}
\caption{Weighted number of galaxies $N$ and average redshift $\la z\ra$ in the KiDS+VIKING-450 survey \citep[KV450,][]{Wright:2018} in pointings of different depth for each of the five tomographic bins used in H18. Each colour corresponds to one redshift bin of H18. A single point represents one percentile of the respective redshift bin, where the fainter points denote pointings of shallower depth.}
\label{fig:nz_of_meanz}
\end{figure}
The previously derived analytic model describes how varying depth between pointings modifies the correlation function due to the correlation between number density and average redshift of source galaxies. While this model serves as an intuitive first approximation, it completely ignores any effects from the large scale structure (LSS) between the closest and the most distant galaxy. Therefore, we do not expect this model to yield accurate, quantitative results for cosmic shear surveys.

@@

An inspection of KiDS-data showed that the redshift distribution of sources is highly correlated with the depth in the $r$-band. We thus chose to separate the survey into 10 percentiles, sorted by $r$-band depth, i.e. if a pointing had a shallower depth than 90\% of the other pointings, it would belong to the first percentile, and so on. For each percentile $m$ and each tomographic redshift bin $i$ we can extract a weighted number of galaxies $N^i_m$ and a source redshift distribution $p^i_m(z)$ following the direct spectroscopic calibration method of H18\footnote{Instead of using the actual number of galaxies, we take the sum of their lensing weights, following H17. Due to this we account for different weighting of galaxies in the shear correlation functions as well as in the average redshift distribution.}. In Fig.~\ref{fig:nz_of_meanz} the average redshift and weighted number of galaxies are plotted for each percentile of each redshift bin, whereas a selection of source redshift distributions is depicted in Fig.~\ref{fig:nofz_percentiles}.
%
\begin{figure}
\centering
\includegraphics[trim = {1cm 0 1.2cm 0.1cm}, clip, width=\linewidth]{images/dist_z_all.pdf}
\caption{The source redshift distributions $p^i_m(z)$ for the best, average, and worst percentile in each tomographic redshift bin.}
\label{fig:nofz_percentiles}
\end{figure}
% Asserting an arbitrary cosmological model, we can transform the source redshift distributions $p^i_m(z)$ into comoving distance probability distributions $\mathcal{L}^i_m(\chi)$.

Given two comoving distance probability distributions of sources, $\mathcal{L}^i_m(\chi)$ and $\mathcal{L}^j_n(\chi)$, we can compute the shear correlation functions from the underlying matter power spectrum, $P_\delta(k,\chi)$, via \citep{1992ApJ...388..272K} \begin{align}
\label{eq:xipm-pkappa}
\xi_{\pm,mn}^{ij}(\theta) =& \int_0^\infty \frac{{\rm d}\ell\,\ell}{2\pi}\, J_{0,4}(\ell\theta)\, P^{ij}_{\kappa,mn}(\ell)\, , \\
\label{eq:pkappa-pdelta/lenseff}
P^{ij}_{\kappa,mn}(\ell) =& \frac{9 H_0^4\Omega_{\rm m}^2}{4c^4}\int_0^{\chi_{\rm{H}}} {\rm d}\chi\, \frac{g^i_m(\chi)g^j_n(\chi)}{a^2(\chi)}\, P_\delta\left(\frac{\ell}{f_K(\chi)},\chi\right)\, , \\
\label{eq:lenseff}
g^i_m(\chi) =& \int_\chi^{\chi_{\rm{H}}} {\rm d}\chi' \, \mathcal{L}^i_m(\chi') \, \frac{f_K(\chi'-\chi)}{f_K(\chi')}\, .
\end{align}
Here, $J_n$ denote the $n$-th order Bessel Functions, $f_K(\chi)$ is the comoving angular diameter distance and $\chi_{\rm{H}}$ is the comoving distance to the horizon. The parameters $H_0$ and $c$ denote the Hubble constant and the speed of light.
